---
title: "Reproducible Research Peer Assignment 1"
author: "Liu, Rong"
date: "Saturday, January 17, 2015"
output: html_document
---

This document is an assignment from Reproducible Research. For more details on the requirements please see
<https://class.coursera.org/repdata-010/human_grading/view/courses/973511/assessments/3/submissions>.

The data for this assignment can be downloaded from the course web site:
<https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip> [52K]

The variables included in this dataset are:
steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

##Loading and preprocessing the data
```{r}
data = read.csv("./activity.csv")
#data[["date"]] = strptime(data[["date"]], "%Y-%m-%d")
summary(data)
```

##What is mean total number of steps taken per day?
```{r message = FALSE}
library(dplyr)
library(ggplot2)
data_day = summarize(group_by(data, date), steps=sum(steps))
data_day = data_day[complete.cases(data_day), ]
ggplot(data_day, aes(x=steps)) + geom_histogram() + ggtitle("Histogram of Steps Taken per Day")
```

Mean steps taken per day: 
```{r}
mean_val = mean(data_day$steps); mean_val
```

Median steps taken per day: 
```{r}
median_val = median(data_day$steps); median_val
```

##What is the average daily activity pattern?
```{r message = FALSE}
data_interval = summarize(group_by(data, interval), steps=mean(steps, na.rm=TRUE))
data_interval = data_interval[complete.cases(data_interval), ]
ggplot(data_interval, aes(y=steps, x=interval)) + geom_line() + ggtitle("Average Number of Steps Taken per Interval")
```

The 5-minute interval that contains the maximum number of steps is:
```{r message = FALSE}
index = which.max(data_interval$steps)
data_interval$steps[index]
```

#Imputing missing values
The total number of missing values in the dataset is:
```{r}
nrow(data) - sum(complete.cases(data))
```

To fill the missing value in any 5-minute interval, We use the mean value for that 5-minute interval across all days. The data after filling missing values is:
```{r}
data_filled = data
for (i in 1:nrow(data)) {
    steps = data$steps[i]
    interval = data$interval[i]
    if (is.na(steps)) {
        data_filled[i, 1] = data_interval[data_interval$interval==interval, 2]
    }
}
summary(data_filled)
```

```{r message=FALSE}
data_filled_day = summarize(group_by(data_filled, date), steps=sum(steps))
ggplot(data_filled_day, aes(x=steps)) + geom_histogram() + ggtitle("Histogram of Steps Taken Per Day (after NA replaced)")
```

The mean of total number of steps taken per day is:
```{r message=FALSE}
mean_val_filled = mean(data_filled_day$steps); mean_val_filled
```

The median of total number of steps taken per day is:
```{r message=FALSE}
median_val_filled = median(data_filled_day$steps); median_val_filled
```

The mean value is as same as that of the orginal value.
The difference between the median value and that of the original data is negligible.

By imputting missing values, variance of the steps taken per day decreased from `r var(data_day$steps)` to `r var(data_filled_day$steps)`

##Are there differences in activity patterns between weekdays and weekends?
```{r message=FALSE}
data_weekday <-
    data_filled %>%
    mutate(day = ifelse(weekdays(as.POSIXct(date)) %in% c("Saturday", "Sunday"), "weekend", "weekday")) %>%
    group_by(interval, day) %>%
    summarize(steps=mean(steps))
ggplot(data_weekday, aes(interval,steps)) + geom_line() + facet_wrap(~day, nrow=2)
```
